Limb-darkening and exoplanets II: choosing the right law for minimum bias retrieval of transit parameters from transit lightcurves

### 2015-10-20 which_law_2_use.tex
Added discussion and conclusions. Changed a line in the introduction.

Given the above, unless the data quality is really poor, there is actually no good reason why one would actually want to fix the 
limb-darkening coefficients in the transit fitting process (assuming computational resources are not an issue). However, although fitting the 
limb-darkening coefficients seems to be a good solution to the accuracy problem, this still has the issue of the low flexibility of the quadratic limb-darkening 
law which still causes important biases on the retrieved transit parameters according to EJ15, which can be a large as $\sim 1\%$ for the planet-to-star 
radius ratio $R_p/R_*$, $\sim 2\%$ for the scaled semi-major axis, $a/R_*$, and $2\%$ for the inclination, where $R_p$ is the planetary radius, $R_*$ 
the stellar radius and $a$ the semi-major axis. The importance for missions like Kepler and for future missions like the Transiting Exoplanet Survey 
Satellite \cite[TESS,][]{ricker2014} is evident when looking at the most recent results from the Kepler mission: if we focus on the planet-to-star radius ratio 
alone, which according to EJ15 has an accuracy bias on the order of $\sim 0.2\%$, a query to the Nasa Exoplanet Archive\footnote{\url{http://exoplanetarchive.ipac.caltech.edu/}; query done on 29/09/2015.} shows that out of 3063 planetary candidates, 933 ($26\%$) have precisions better than 
$0.2\%$, and out of 1001 Kepler confirmed exoplanets, 463 (46\%) do too. This means that at least\footnote{It is important to note that this is an 
underestimate of the number of planet candidates with estimation errors, due to the fact that the Kepler pipeline makes use of fixed limb-darkening 
coefficients using the quadratic law \citep{rowe2015}, which as explained here and estimated in EJ15, gives rise to more severe biases than the one used for this 
estimation} $26\%$ of the planet candidates (from which, e.g., population studies, which rely on averaging out \textit{random} and not \textit{systematic} 
uncertainties like the ones introduced by limb-darkening) and almost half of the confirmed exoplanets (from which, e.g., characterization studies are 
based on) have important systematic errors.

@@

This work is organised as follows. In \S2, we first revisit the logarithmic and exponential 
laws in order to try to derive an efficient informative sampling strategy following the methods in \cite{kipping2013} for the standard forms of these laws. 
In \S3 we use those results in addition to the ones published by \cite{kipping2013} and \cite{kipping2015} in order to simulate transit lightcurves to test 
how well these laws perform in retrieving transit parameters. In \S4 we present a discussion of our work and present our conclusions in \S5.

@@

Overall, as expected, the law that retrieves the parameters with minimum bias in low impact-parameter transits is the three-parameter law. However, for 
high impact-parameter transits of small planets, the logarithmic and square-root law seem to outperform this law in general if the objective is to retrieve 
a minimum bias estimator for $p$. This is probably due to the fact that the law is so flexible that it allows for small distortions of the transit lightcurve that, 
in this case, mimics variations that were actually attributable to $p$, adding a small but important bias on the retrieved parameter (\textbf{NE: maybe I should 
plot a transit fit showing this?}). However, for the other 
parameters ($a_R$ and $i$) this law seems to be the best choice in terms of the achieved accuracy.

\section{Discussion}
\section{Conclusions}

>>>>

Given the above, unless the data quality is really poor (and one is willing to sacrifice bias over variance), there is actually no good reason why one would 
actually want to fix the limb-darkening coefficients in the transit fitting process (assuming computational resources are not an issue). However, although fitting the 
limb-darkening coefficients seems to be a good solution to the accuracy problem, this still has the issue of the low flexibility of the quadratic limb-darkening 
law which still causes important biases on the retrieved transit parameters according to EJ15, which can be a large as $\sim 1\%$ for the planet-to-star 
radius ratio $R_p/R_*$, $\sim 2\%$ for the scaled semi-major axis, $a/R_*$, and $2\%$ for the inclination, where $R_p$ is the planetary radius, $R_*$ 
the stellar radius and $a$ the semi-major axis. The importance for missions like Kepler and for future missions like the Transiting Exoplanet Survey 
Satellite \cite[TESS,][]{ricker2014} is evident when looking at the most recent results from the Kepler mission: if we focus on the planet-to-star radius ratio 
alone, which according to EJ15 has an accuracy bias on the order of $\sim 0.2\%$, a query to the Nasa Exoplanet Archive\footnote{\url{http://exoplanetarchive.ipac.caltech.edu/}; query done on 29/09/2015.} shows that out of 3063 planetary candidates, 933 ($26\%$) have precisions better than 
$0.2\%$, and out of 1001 Kepler confirmed exoplanets, 463 (46\%) do too. This means that at least\footnote{It is important to note that this is an 
underestimate of the number of planet candidates with estimation errors, due to the fact that the Kepler pipeline makes use of fixed limb-darkening 
coefficients using the quadratic law \citep{rowe2015}, which as explained here and estimated in EJ15, gives rise to more severe biases than the one used for this 
estimation} $26\%$ of the planet candidates (from which, e.g., population studies, which rely on averaging out \textit{random} and not \textit{systematic} 
uncertainties like the ones introduced by limb-darkening) and almost half of the confirmed exoplanets (from which, e.g., characterization studies are 
based on) have important systematic errors.

@@

This work is organised as follows. In \S2, we first revisit the logarithmic and exponential 
laws in order to try to derive an efficient informative sampling strategy following the methods in \cite{kipping2013} for the standard forms of these laws. 
In \S3 we use those results in addition to the ones published by \cite{kipping2013} and \cite{kipping2015} in order to simulate transit lightcurves to test 
how well these laws perform in retrieving transit parameters. In \S4 we present a discussion and conclusions of our work.

@@

Overall, as expected, the law that retrieves the parameters with minimum bias in low impact-parameter transits is the three-parameter law. However, for 
high impact-parameter transits of small planets, the logarithmic and square-root law seem to outperform this law in general if the objective is to retrieve 
a minimum bias estimator for $p$. This is due to the fact that the law is so flexible that it allows for small distortions of the transit lightcurve that, 
in this case, mimics variations that were actually attributable to $p$, adding a small but important bias on the retrieved parameter. However, for the other 
parameters ($a_R$ and $i$) this law seems to be the best choice in terms of the achieved accuracy.

\section{Discussion \& Conclusions}
In this work we have explored the different biases on the retrieved transit parameters by the usage of different limb-darkening laws. In particular, we 
explored the biases introduced by using three two-parameter laws (the quadratic, logarithmic and square-root laws) and the three-parameter law, for which 
we showed that a careful selection of the limb-darkening law to be used in a given dataset can allow one to retrieve minimum bias estimators for the transit 
parameters from transit lightcurves. In general, as shown in \S3, these two-parameter laws have ranges in which they perform at the same level or even better 
than the three-parameter law and, because of this, we encourage their use. We also discourage the usage of the exponential law which, as shown in \S2.2 is 
unphysical, and can lead to numerical errors if used on real data where sampling of points close to the limb (where the law diverges) is not uncommon.

The question of \textit{when} to use each of those laws is an important one. In particular, we have explored the relation between the \textit{real}, underlying 
transit parameters, the stellar host properties and the bias on the retrieved transit parameters, but we have stated no specific rule regarding when to use each 
law, but only guidelines based on our simulations. This was made on purpose, as the selection of which law to use has to be done in a case-by-case basis, reason 
by which we provide a code with which one can obtain the results for the same simulations for specific 
cases and with different resolutions\footnote{\url{http://www.github.com/nespinoza/ld-simulations/}}. It is important to note that in this work we have used the 
Kepler bandpass to study the relation between the biases on the retrieved transit parameters and the stellar host properties, but different bandpasses should give 
rise to slightly different functional relations because they give rise to different model intensity profiles, which are used in those simulations to approximate the 
expected intensity profile and the adequacy of each limb-darkening law at following it. As such, limb-darkening coefficients have to be calculated for each 
particular case, which is straightforward to do using the algorithms described in EJ15\footnote{\url{http://www.github.com/nespinoza/limb-darkening/}}.
